---
title: "Week 1"
description: |
  Overview of Multilevel Models
output: 
  distill::distill_article:
    toc: true
---

```{r week-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("helper.R")
```

Before you start the materials for week 1, make sure you have reviewed the [course overview video](../) and the [syllabus](../syllabus). 

# Week Learning Objectives

By the end of this module, you will be able to 

- Identify alternative names for multilevel modeling (MLM)
- Describe the types of data MLM can handle
- Knit a simple R Markdown file

## Task Lists

1. Install/Update R and RStudio on your computer
2. Review the resources (lecture videos and slides)
3. Complete the assigned readings
    * Snijders & Bosker chapters 1 and [2][4] (they're relatively short)
    * [R Markdown Intro][1]
    * [Markdown Quick Reference][2]
4. Introduce yourself on the #introduce-ourselves Slack channel (as part of [HW 1](../homework-1))
5. Attend the synchronous Zoom session to learn more about R Markdown
6. Complete [Homework 1](../homework-1)

[1]: https://vimeo.com/178485416
[2]: https://web.mit.edu/r/current/RStudio/resources/markdown_help.html
[4]: https://us.sagepub.com/sites/default/files/upm-assets/44075_book_item_44075.pdf

# Lecture

```{r, echo=FALSE}
add_youtube("ufvCUaUkYOA")
```

```{r, include=FALSE}
library(checkdown)
right_ans <- "Correct :+1:"
wrong_ans <- "That is not correct. Rewatch the video if needed"
```

***

Check your learning: How would you describe the data structure in the video?

```{r echo=FALSE, results="asis"}
check_question("hierarchical", 
               options = c("observational", "educational", "hierarchical"), 
               type = "radio", 
               right = right_ans, 
               wrong = wrong_ans, 
               alignment = TRUE)
```

***

```{r, echo=FALSE}
add_youtube("9Ax8W6VPgmk")
```

***

Check your learning: What is another name for multilevel modeling? 

```{r echo=FALSE, results="asis"}
check_question(c("mixed-effect modeling"),  
               options = c("mixed-effect modeling", "hierarchical regression", 
                           "fixed-effect modeling"), 
               type = "radio", 
               right = right_ans, 
               wrong = wrong_ans)
```

***

```{r, echo=FALSE}
add_youtube("4fl2chIkDy4")
```

***

Check your learning: In the data structure where there are multiple measurements for the same person, which level is level 1? 

```{r echo=FALSE, results="asis"}
check_question(c("measurement"),  
               options = c("measurement", "person"), 
               type = "radio", 
               right = right_ans, 
               wrong = wrong_ans, 
               alignment = TRUE)
```

***

# R Codes

```{r, echo=FALSE}
add_youtube("G-IjAC-2uik")
```

```{r, echo=FALSE}
add_youtube("-PdycAVwMuY")
```

```{r r-code, child='rcodes/week1-intro-RMarkdown.RMarkdown'}

```