---
title: "Week 7"
description: |
  Multilevel Models for Experimental Data
output: 
  distill::distill_article:
    toc: true
params:
  slides: "07_mlm_for_experimental_data"
---


```{r week-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("helper.R")
```

# Week Learning Objectives

By the end of this module, you will be able to 

- Identify the correct levels with experimental studies
- Describe designs with crossed random levels
- Assign variables to appropriate levels, and tell which variables can have random slopes at which levels
- Compute a version of effect size ($R^2$) for experimental data

Note: In the video the slides use $d$ for effect size, but we will use $R^2$ in this class.

## Task Lists

1. Review the resources (lecture videos and slides)
2. Complete the assigned readings
    * Snijders & Bosker 13.1
    * [Hoffman & Rovine (2007)](https://link.springer.com/content/pdf/10.3758/BF03192848.pdf)
3. Attend the Thursday session and participate in the class exercise
4. Complete Homework 7
5. Think about your project---Prospectus will be due in two weeks
6. Additional resources for learning MLM for experimental designs
    * [This paper](https://www.annualreviews.org/doi/abs/10.1146/annurev-psych-122414-033702) by Judd et al. 

# Lecture

## Slides

You can view and download the slides here: 
[PDF](`r file.path("slides", paste0(params$slides, ".pdf"))`){target="_blank"}

```{r, include = FALSE}
library(checkdown)
right_ans <- c("You got it :+1:", 
               "Great job :+1:")
wrong_ans <- "Try again. You can do it!"
```

## Multilevel Experiments

```{r, echo=FALSE}
add_youtube("9btZ_FrDGus")
```

***

Check your learning: In a research study, 10 hospitals are randomly assigned to a treatment condition to adopt a new drug, whereas the other 10 hospitals use the conventional method. What is the design of this study?

```{r echo=FALSE, results="asis"}
check_question("A cluster-randomized trial", 
               options = c("A cluster-randomized trial", 
                           "A multisite trial", 
                           "A within-subjects design"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

## Example

```{r, echo=FALSE}
add_youtube("4baFBogxmAs")
```

***

Check your learning: In the data set, how many observations are there at level 1?

```{r echo=FALSE, results="asis"}
check_question("7,803", 
               options = c("7,803", 
                           "153", 
                           "51"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans, 
               alignment = TRUE)
```

***

## Long vs. Wide data set

```{r, echo=FALSE}
add_youtube("NsRnhNz3-10")
```

***

Check your learning: In the following data, `hvltt` to `hvltt4` are the test score of a verbal learning test across four time points. Is this a long or a wide data set?

```{r, message=FALSE}
library(tidyverse)
read.csv("https://books.psychstat.org/rdata/data/active.csv",
         header = TRUE) %>%
  select(id, hvltt:hvltt4) %>%
  rmarkdown::paged_table()
```


```{r echo=FALSE, results="asis"}
check_question("wide", 
               options = c("long", 
                           "wide"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans, 
               alignment = TRUE)
```

## Crossed Random Levels

```{r, echo=FALSE}
add_youtube("Vgv78TgDWgA")
```

***

Think more: What is the data structure if there are 1,000 students from 100 schools and 30 neighborhoods, and each school has students from multiple neighborhoods?

***

### Unconditional model

```{r, echo=FALSE}
add_youtube("zECzzA4H_uo")
```

***

Practice yourself: Compute the design effects for the person level and for the item level. Do the design effects suggest the need for both levels? (That is, are both design effects > 1.1?)

Answer: See the computation in the [R code](rcode7.html#intraclass-correlations-and-design-effects)

***

## Cross-Classified Random-Effect Model With Random Slopes

```{r, echo=FALSE}
add_youtube("dyDKHC0sugE")
```

***

Check your learning: If in the experiment, each person respond to each item 3 times, each time with a different `duration`. At what level(s) can `duration` have random slopes?

```{r echo=FALSE, results="asis"}
check_question("Both person and item", 
               options = c("Both person and item", 
                           "Person only", 
                           "Item only", 
                           "Within-cell"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

## Full Model

```{r, echo=FALSE}
add_youtube("bIrEtjB91gk")
```

## Effect Size

Please check out the slides and the examples in the [R code](rcode7.html#effect-size)
