---
title: "Week 10"
description: |
  Models for Longitudinal Data II
output: 
  distill::distill_article:
    toc: true
params:
  slides: "10_longitudinal_analysis_2"
---


```{r week-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("helper.R")
```

# Week Learning Objectives

By the end of this module, you will be able to 

- Specify models with alternative error covariance structures
- Describe the difference between analyzing trends vs. analyzing **fluctuations** with longitudinal data
- Run analyses with **time-varying predictors** (i.e., level-1 predictors)
- Interpret and plot results

## Task Lists

1. Review the resources (lecture videos and slides)
2. Complete the assigned readings
    * [Hoffman (2014) ch 4.1](https://www-taylorfrancis-com.libproxy1.usc.edu/books/9781315744094/chapters/10.4324/9781315744094-12) (USC SSO required)
    * [Hoffman (2014) ch 8](https://www-taylorfrancis-com.libproxy1.usc.edu/books/9781315744094/chapters/10.4324/9781315744094-17) (USC SSO required)
3. Attend the Thursday session and participate in the class exercise
4. Complete Homework 8
5. (Optional) Read the bonus R code on the generalized estimating equations (GEE) method
    * Additional reference: https://journals.sagepub.com/doi/abs/10.3102/10769986211017480

# Lecture

## Slides

<span style="color:red">Note that in some of the videos below the Bayeisan analyses were used; however, for the class this year we will stay with frequentist analyses. The results and interpretations are basically; just note some differences in the terminology.</span>

You can view and download the slides here: 
[PDF](`r file.path("slides", paste0(params$slides, ".pdf"))`){target="_blank"}

```{r, include=FALSE}
library(checkdown)
right_ans <- c("Correct :+1:", 
               "Great job :+1:")
wrong_ans <- "Try again. You can do it!"
```

## Longitudinal Data Analysis II

```{r, echo=FALSE}
add_youtube("k5WyosL97B0")
```

## Temporal Covariance/Correlation

```{r, echo = FALSE}
add_youtube("pzsL1y5XsQ0")
```

Check your learning: Assume that the temporal correlation decreases with a longer time gap. A researcher collects data at baseline (Time 1), 3-month follow-up (Time 2), and then 5-month follow-up (Time 3). Which correlation should be strongest?

```{r echo = FALSE, results = "asis"}
check_question("Between Time 2 and Time 3",
               options = c("Between Time 1 and Time 2",
                           "Between Time 2 and Time 3",
                           "Between Time 1 and Time 3"),
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = "Think about which time period is shortest")
```

### Covariance Structure in MLM

OLS and RI-MLM/RM-ANOVA

```{r, echo = FALSE}
add_youtube("UKxumumURd0")
```

Check your learning: The random-intercept model/repeated-measures ANOVA assumes a specific temporal covariance structure. What is that structure called?

```{r echo = FALSE, results = "asis"}
check_question("Compound Symmetry",
               options = c("Independence",
                           "Autoregressive",
                           "Autoregressive and Moving Average",
                           "Compound Symmetry"),
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

Random Slopes

```{r, echo = FALSE}
add_youtube("UiBMcbNGbHk")
```

Autoregressive(1) error structure

```{r, echo = FALSE}
add_youtube("Z5ZghnvZkYI")
```

Check your learning: In an AR(1) covariance structure, what is the implied correlation between Time 2 and Time 4, if $\rho = .4$?

```{r echo = FALSE, results = "asis"}
check_question(".16",
               options = c(".4",
                           ".2",
                           ".16",
                           ".064"),
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans, 
               alignment = TRUE)
```

## Analyzing Dynamics

```{r, echo=FALSE}
add_youtube("xW_84TYK85g?start=105")
```

```{r, echo=FALSE}
add_youtube("F-NNPPAuPWE")
```

***

Check your learning: When analyzing a conversation between a couple, a researcher is interested in whether a person follow up the partner's complaints with positive or negative behaviors. Is this an example of studying trends or fluctuations?

```{r echo=FALSE, results="asis"}
check_question("Fluctuations", 
               options = c("Trends", 
                           "Fluctuations"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans, 
               alignment = TRUE)
```

***

## Model 1

```{r, echo = FALSE}
add_youtube("mDdyowlEUic")
```

***

Check your learning: In the model above, what is the interpretation of the contextual effect of `mood1`?

```{r echo=FALSE, results="asis"}
check_question("The predicted difference in symptoms between two individuals with 1 unit difference in average negative mood on a day when both individuals have the same daily mood level", 
               options = c("The predicted difference in average symptoms between two individuals with 1 unit difference in average negative mood", 
                           "The predicted difference in symptoms for an individual across two days with 1 unit difference in mood", 
                           "The predicted difference in symptoms between two individuals with 1 unit difference in average negative mood on a day when both individuals have the same daily mood level"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

## Model 2

```{r, echo=FALSE}
add_youtube("bbNaNi2hUXQ")
```

Note: For the coefficients of `stressor` and `stressor_pm` in the above model, the coefficients are ones adjusting for the other predictors in the model (e.g., `mood1_pm`, `mood1_pmc`, `women`, and their interactions). 
