---
title: "Week 8"
description: |
  Models for Longitudinal Data I
output: 
  distill::distill_article:
    toc: true
params:
  slides: "08_longitudinal_analysis_1"
---


```{r week-setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
source("helper.R")
```

# Week Learning Objectives

By the end of this module, you will be able to 

- Describe the similarities and differences between longitudinal data and cross-sectional clustered data
- Perform some basic **attrition analyses**
- Specify and run **growth curve analysis**
- Analyze models with **time-invariant covariates** (i.e., lv-2 predictors) and interpret the results

Note: In the video the slides use $d$ for effect size, but we will use $R^2$ in this class.

## Task Lists

1. Review the resources (lecture videos and slides)
2. Complete the assigned readings
    * Snijders & Bosker ch 15 (you can skip 15.1.3 and 15.1.4)
3. Prepare the prospectus and schedule a meeting with the instructor
4. Additional resources for learning MLM for longitudinal data analysis
    * [This excellent book by Hoffman (2014)](https://www-taylorfrancis-com.libproxy2.usc.edu/books/9781315744094) (USC SSO required)

# Lecture

## Slides

<span style="color:red">Note that in some of the videos below the Bayeisan analyses were used; however, for the class this year we will stay with frequentist analyses. The results and interpretations are basically; just note some differences in the terminology.</span>

You can view and download the slides here: 
[PDF](`r file.path("slides", paste0(params$slides, ".pdf"))`){target="_blank"}

```{r, include=FALSE}
library(checkdown)
right_ans <- c("Correct :+1:", 
               "Great job :+1:")
wrong_ans <- "Try again. You can do it!"
```

## Longitudinal Data Analysis

```{r, echo = FALSE}
add_youtube("bdDrcmktBnE")
```

***

Check your learning: In a research study, data were collected for a group of patients on symptoms of eating disorder on a weekly interval across 5 weeks. What type of data is this?

```{r echo = FALSE, results = "asis"}
check_question("Panel data", 
               options = c("Panel data", 
                           "Intensive longitudinal data", 
                           "Cross-sectional data"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

## Example data

```{r, echo = FALSE}
add_youtube("3PDA_ur1sOU")
```

***

Check your learning: In the data set, at what level is `homecog`, which is a measure of mother's cognitive stimulation at baseline?

```{r echo = FALSE, results = "asis"}
check_question("Level 2", 
               options = c("Level 1", 
                           "Level 2"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans, 
               alignment = TRUE)
```

***

### Basic attrition analysis

```{r, echo = FALSE}
add_youtube("b2dto5JwRQ8")
```

See the [R code section](../rcode8/#attrition-analyses). 

***

Check your learning: In the spaghetti plot, what does the average trend line mean?

```{r echo = FALSE, results = "asis"}
check_question("That the sample average reading score increased across time", 
               options = c("That every child showed an increasing trend", 
                           "That the sample average reading score increased across time", 
                           "That there is no variability in individual trajectories"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

## Growth Curve Modeling

<span style="color:red">Note that in the video, the function `brm()` from the `brms` package was used for Bayesian analyses. However, in this class we will use the `glmmTMB()` function instead (see the updated slides). The parameter estimates and interpretations are basically the same.</span>

```{r, echo = FALSE}
add_youtube("iT7kzf2gIUs")
```

***

Thinking exercise: In a growth model, what does it mean when $\tau_1 = 0$?

```{r echo = FALSE, results = "asis"}
check_question("That every participant has the same rate of change", 
               options = c("That on average, there is no change on the outcome across time", 
                           "That every participant has the same rate of change", 
                           "That every participant has the same outcome score"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

### Linear growth

<span style="color:red">Note that what is labelled as SD~post~ is the Bayesian analogue of the standard error.</span>

```{r, echo = FALSE}
add_youtube("4xHz4OIPVh8")
```

***

Check you learning: What is the advantage of having time to start at 0?

```{r echo = FALSE, results = "asis"}
check_question("The intercept corresponds to the predicted score at the first time point", 
               options = c("The intercept corresponds to the predicted score at the first time point", 
                           "So that tau_1 will be the SD of the slopes"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans)
```

***

### Piecewise linear growth

```{r, echo = FALSE}
add_youtube("p6TQdUzi1d0")
```

***

Practice yourself: What should the coding of phase 1 and phase 2 be if the turning point is set at `time` = 2?

***

<span style="color:red">Instead of using the LOOIC in the Bayesian analysis as discussed in the video, we can use the more popular AIC statistic to compare the two models.</span>

```{r, echo = FALSE}
add_youtube("MbDcv05wOrU")
```

Note: In this example, the turning point was chosen mostly based on the spaghetti plot and was arbitrary. For your research, you should justify your choice. 

***

Check your learning: If a piecewise growth model has an AIC of 23745, and a linear growth model has an AIC of 23650, which model should be preferred?

```{r echo = FALSE, results = "asis"}
check_question("Linear",
               options = c("Piecewise", 
                           "Linear"), 
               type = "radio", 
               right = sample(right_ans, 1, prob = c(.7, .3)), 
               wrong = wrong_ans, 
               alignment = TRUE)
```

***

## Time-Invariant Covariates 

```{r, echo = FALSE}
add_youtube("fI6XJnVQ3pg")
```

***

Thinking exercise: What does the coefficient for `phase1` mean when the model includes an interaction between `phase1` and `homecog9`?

```{r echo = FALSE, results = "asis"}
check_question("The average slope from time 0 to time 1 for kids with homecog = 9",
               options = c("The average slope from time 0 to time 1",
                           "The average slope from time 0 to time 1 for kids with homecog = 9",
                           "The average slope from time 0 to time 1, holding constant homecog"),
               type = "radio",
               right = sample(right_ans, 1, prob = c(.7, .3)),
               wrong = wrong_ans)
```
